\speculake: Hiding Malware in Speculative Execution

### 2018-12-07 paper/architecture.tex
add table
add reference to table 1

We observe that Intel CPUs consider three types of x86\_64 indirect branches:
\texttt{retq}, \texttt{callq *\%rax}, and \texttt{jmpq *\%rax}\footnote{other
general purpose registers besides \texttt{\%rax} can be used as well}.
We created a simple trigger program that performs a series of 28 indirect
branches
using \texttt{jmpq *\%rax} instructions, as we found that 28 jumps was the minimum
needed to reliably train the indirect branch predictor. 
Between each jump, we incremented
\texttt{\%rax} accordingly to continue on to the next jump. After these 28
jumps, we load a function pointer into \texttt{\%rax} and do a final indirect
branch using \texttt{callq *\%rax}. In our trigger, we perform these jumps
        repeatedly in a loop.

In our payload program, we first perform the same 28 indirect jumps. We ensure
the source and destination addresses of these jumps is the same as in the trigger program by
manually defining their containing function at a fixed address inside a linker
script. We also do the final indirect call to a function pointer, but with two
differences. First, the destination in the function pointer is a different
address, and second, the memory location of the function pointer itself is uncached.
This forces the CPU to predict the destination of the final indirect call while it
waits for the function pointer to load from memory. Due to the similar
history of branches with the trigger program,
the CPU will (incorrectly) predict the destination to be the same as the one in
the trigger program, which determines the speculative entry point for the
payload. Even though the in-order execution of
payload program never executes or even reads from this address, the CPU will
briefly execute instructions there speculatively.

>>>>

We observe that Intel CPUs consider three types of x86\_64 indirect branches:
\texttt{retq}, \texttt{callq *\%rax}, and \texttt{jmpq *\%rax}\footnote{other
general purpose registers besides \texttt{\%rax} can be used as well}. We
created a simple trigger program that performs a series of indirect branches
using \texttt{jmpq *\%rax} instructions.
%, as we found that 28 jumps was the
%minimum needed to reliably train the indirect branch predictor. 
Between each
jump, we incremented \texttt{\%rax} accordingly to continue on to the next jump.
After these jumps, we load a function pointer into \texttt{\%rax} and do a
final indirect branch using \texttt{callq *\%rax}. In our trigger, we perform
these jumps repeatedly in a loop.

In our payload program, we first perform the same indirect jumps. We ensure
the source and destination addresses of these jumps is the same as in the
trigger program by manually defining their containing function at a fixed
address inside a linker script. We also do the final indirect call to a function
pointer, but with two differences. First, the destination in the function
pointer is a different address, and second, the memory location of the function
pointer itself is uncached. This forces the CPU to predict the destination of
the final indirect call while it waits for the function pointer to load from
memory. Due to the similar history of branches with the trigger program, the CPU
will (incorrectly) predict the destination to be the same as the one in the
trigger program, which determines the speculative entry point for the payload.
Even though the in-order execution of payload program never executes or even
reads from this address, the CPU will briefly execute instructions there
speculatively. \processorTable

In Table~\ref{table:processors} we analyze the number of necessary training
indirect jumps various processors require to consistently ($> 95\%$) have the
payload program enter the speculative world at the location primed in the
trigger program. We found that 28 indirect jumps was sufficient for our trigger
program on each of the test processors to reliably ensure the speculative
execution began at the correct address.
